---
title: "Build confounders dataset"
author: "Derek Burk"
date: "9/24/2021"
output: html_document
---


# Packages

```{r}
library(dplyr)
library(readr)
library(rlang)
library(gt)
```


# File paths

These file paths have to start with ".." because the working directory is set 
to the directory in which this RMarkdown file is saved 
("scripts_that_manipulate_downloaded_data").

```{r}
countrylist_path <- "../countrylist.csv"
age_dist_path <- "../downloaded_data/age-dist-2021-07-16.csv"
gdp_path <- "../downloaded_data/gdp-growth-world-bank-2021-09-24.csv"
gini_path <- "../downloaded_data/gini-world-bank-2021-09-24.csv"
life_expectancy_path <- 
  "../downloaded_data/life-expectancy-world-bank-2021-09-24.csv"
maternal_mortality_path <- 
  "../downloaded_data/maternal-mortality-world-bank-2021-09-24.csv"
poverty_path <- "downloaded_data/poverty-world-bank-2021-07-15.csv"
under_5_mortality_path <- 
  "../downloaded_data/under-5-mortality-world-bank-2021-07-16.csv"
```


# Country list

```{r}
countrylist <- read_csv(
  countrylist_path,
  col_names = c("row_num", "country"),
  col_types = cols(row_num = col_skip(), country = col_character()),
  skip = 1
)
```


# Function to match country names

```{r}
match_country_names <- function(country_names) {
  
  no_match <- sort(setdiff(countrylist$country, country_names))
  
  if (length(no_match) == 0) {
    cat("All countries matched!")
  } else {
    warning(
      paste0("No match for:\n\n", paste0(no_match, collapse = "\n")),
      call. = FALSE
    )
  }
  
}
```



# Age distribution

```{r}
age_dist <- read_csv(
  age_dist_path,
  col_types = cols(
    LocID = col_double(),
    Location = col_character(),
    VarID = col_double(),
    Variant = col_character(),
    Time = col_double(),
    MidPeriod = col_double(),
    AgeGrp = col_character(),
    AgeGrpStart = col_double(),
    AgeGrpSpan = col_double(),
    PopMale = col_double(),
    PopFemale = col_double(),
    PopTotal = col_double()
  )
)
```

## Reconcile country names

```{r}
match_country_names(age_dist$Location)
```


```{r}
age_dist <- age_dist %>% 
  rename(country = Location) %>% 
  mutate(
    country = recode(
      country,
      "United Kingdom" = "Britain",
      "Czechia" = "Czech Republic",
      "Dem. People's Republic of Korea" = "South Korea",
      "United States of America" = "United States"
    )
  )

match_country_names(age_dist$country)
```
## Filter down to countries in analysis

```{r}
age_dist_subset <- age_dist %>% 
  inner_join(countrylist, by = "country")
```

## Filter down to most recent year (no later than 2019):

```{r}
age_dist_subset <- age_dist_subset %>% 
  filter(Time < 2020) %>% 
  group_by(country) %>% 
  mutate(most_recent_year = max(Time, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(Time == most_recent_year)
```

Distribution of most recent year:

```{r}
age_dist_subset %>% 
  distinct(country, most_recent_year) %>% 
  count(most_recent_year)
```


## Create indices

Create percent over 65 and old-age dependency indices:

```{r}
age_indicators <- age_dist_subset %>% 
  rename(year = Time) %>% 
  group_by(country, year) %>% 
  summarize(
    percent_over_65 = 
      100 * sum(PopTotal[AgeGrpStart >= 65]) / sum(PopTotal),
    old_age_dependency = 
      100 * sum(PopTotal[AgeGrpStart >= 65]) / sum(PopTotal[AgeGrpStart < 65]),
    .groups = "drop"
  )

age_indicators %>% 
  arrange(desc(percent_over_65)) %>% 
  mutate(across(c(percent_over_65, old_age_dependency), ~round(.x, 1))) %>% 
  gt() %>% 
  tab_header(
    title = "% over 65 and old-age dependency ratio", 
    subtitle = "Descending order by % over 65"
  )
```


# GDP growth

```{r}
gdp <- read_csv(
  gdp_path,
  col_types = cols(
    iso2c = col_character(),
    country = col_character(),
    NY.GDP.PCAP.KD.ZG = col_double(),
    year = col_double(),
    iso3c = col_character(),
    region = col_character(),
    capital = col_character(),
    longitude = col_double(),
    latitude = col_double(),
    income = col_character(),
    lending = col_character()
  )
)
```

## Reconcile country names

```{r}
match_country_names(gdp$country)
```

```{r}
gdp <- gdp %>% 
  mutate(
    country = recode(
      country,
      "United Kingdom" = "Britain",
      "Slovak Republic" = "Slovakia",
      "Korea, Rep." = "South Korea"
    )
  )

match_country_names(gdp$country)
```

## Filter down to countries in analysis

```{r}
gdp_subset <- gdp %>% 
  inner_join(countrylist, by = "country")
```

## Subset to relevant variables

```{r}
gdp_subset <- gdp_subset %>% 
  select(country, year, gdp_growth_rate = NY.GDP.PCAP.KD.ZG)
```



Not sure if we want to take an average over a certain time period here?


# Gini coefficient

```{r}
gini <- read_csv(
  gini_path,
  col_types = cols(
    iso2c = col_character(),
    country = col_character(),
    SI.POV.GINI = col_double(),
    year = col_double(),
    iso3c = col_character(),
    region = col_character(),
    capital = col_character(),
    longitude = col_double(),
    latitude = col_double(),
    income = col_character(),
    lending = col_character()
  )
)
```


