---
title: "Download life expectancy data from World Bank API"
date: "`r Sys.Date()`"
output: html_document
---

# Packages

```{r}
library(WDI)
library(readr)
```


# Poverty headcount data

From https://data.worldbank.org/indicator/SP.DYN.LE00.IN?view=chart :

( 1 ) United Nations Population Division. World Population Prospects: 2019
Revision, or derived from male and female life expectancy at birth from sources
such as: ( 2 ) Census reports and other statistical publications from national
statistical offices, ( 3 ) Eurostat: Demographic Statistics, ( 4 ) United
Nations Statistical Division. Population and Vital Statistics Reprot ( various
years ), ( 5 ) U.S. Census Bureau: International Database, and ( 6 ) Secretariat
of the Pacific Community: Statistics and Demography Programme.

```{r}
life_expectancy_data <- WDI(
  indicator = "SP.DYN.LE00.IN", 
  start = 2010, 
  end = 2020, 
  extra = TRUE
)
if (!dir.exists("../downloaded_data")) dir.create("../downloaded_data")
write_csv(
  life_expectancy_data, 
  path = paste0(
    "../downloaded_data/life-expectancy-world-bank-", 
    Sys.Date(), 
    ".csv"
  )
)
```

