---
title: "Which countries have data on confounders?"
author: "Derek Burk"
date: "7/16/2021"
output: html_document
---

# Setup

## Packages

```{r}
suppressPackageStartupMessages({
  library(rnaturalearth)
  library(rnaturalearthdata)
  library(ggplot2)
  library(dplyr)
  library(readr)
  library(stringr)
})
```


## File paths

```{r}
data_dir <- "../downloaded_data"
wmd_path <- file.path(data_dir, "wmd-2021-07-15.csv")
poverty_path <- file.path(data_dir, "poverty-world-bank-2021-07-15.csv")
age_path <- file.path(data_dir, "age-dist-2021-07-16.csv")
child_mortality_path <- file.path(
  data_dir, 
  "under-5-mortality-world-bank-2021-07-16.csv"
)
```


# Load data

```{r}
wmd <- read_csv(
  wmd_path,
  col_types = cols(
    country_name = col_character(),
    year = col_double(),
    time = col_double(),
    time_unit = col_character(),
    deaths = col_double()
  )
)
poverty <- read_csv(
  poverty_path,
  col_types = cols(
    iso2c = col_character(),
    country = col_character(),
    SI.POV.DDAY = col_double(),
    year = col_double(),
    iso3c = col_character(),
    region = col_character(),
    capital = col_character(),
    longitude = col_double(),
    latitude = col_double(),
    income = col_character(),
    lending = col_character()
  )
)
age <- read_csv(
  age_path,
  col_types = cols(
    LocID = col_double(),
    Location = col_character(),
    VarID = col_double(),
    Variant = col_character(),
    Time = col_double(),
    MidPeriod = col_double(),
    AgeGrp = col_character(),
    AgeGrpStart = col_double(),
    AgeGrpSpan = col_double(),
    PopMale = col_double(),
    PopFemale = col_double(),
    PopTotal = col_double()
  )
)
child_mortality <- read_csv(
  child_mortality_path,
  col_types = cols(
    iso2c = col_character(),
    country = col_character(),
    SH.DYN.MORT = col_double(),
    year = col_double(),
    iso3c = col_character(),
    region = col_character(),
    capital = col_character(),
    longitude = col_double(),
    latitude = col_double(),
    income = col_character(),
    lending = col_character()
  )
)
```

# Brauner et al. countries

```{r}
brauner_countries <- c(
  "Albania", "Andorra", "Austria", "Belgium", "Bosnia and Herzegovina", 
  "Bulgaria", "Croatia", "Czech Republic", "Denmark", "Estonia", "Finland", 
  "France", "Georgia", "Germany", "Greece", "Hungary", "Iceland", "Ireland", 
  "Israel", "Italy", "Latvia", "Lithuania", "Malaysia", "Malta", "Mexico", 
  "Morocco", "Netherlands", "New Zealand", "Norway", "Poland", "Portugal", 
  "Romania", "Serbia", "Singapore", "Slovakia", "Slovenia", "South Africa", 
  "Spain", "Sweden", "Switzerland", "United Kingdom"
)
```


# Which countries have enough data?

## Mortality

Fix country names to match Brauner:

```{r}
wmd <- wmd %>% 
  mutate(
    country_name = case_when(
      country_name == "Bosnia" ~ "Bosnia and Herzegovina",
      country_name == "Czechia" ~ "Czech Republic",
      TRUE ~ country_name
    )
  )
```

```{r}
countries_with_complete_weekly_data <- wmd %>% 
  filter(time_unit == "weekly" & year %in% 2015:2020) %>% 
  group_by(country_name, year) %>% 
  summarize(has_all_weeks = all(1:52 %in% time), .groups = "drop") %>% 
  group_by(country_name) %>% 
  summarize(for_all_years = all(has_all_weeks), .groups = "drop") %>% 
  filter(for_all_years) %>% 
  pull(country_name) %>% 
  sort()
countries_with_complete_monthly_data <- wmd %>% 
  filter(time_unit == "monthly" & year %in% 2015:2020) %>% 
  group_by(country_name, year) %>% 
  summarize(has_all_months = all(1:12 %in% time), .groups = "drop") %>% 
  group_by(country_name) %>% 
  summarize(for_all_years = all(has_all_months), .groups = "drop") %>% 
  filter(for_all_years) %>% 
  pull(country_name) %>% 
  sort()
countries_with_complete_mortality_data <- c(
  countries_with_complete_weekly_data, countries_with_complete_monthly_data
) %>% 
  unique() %>% 
  sort()
```


## Poverty

```{r}
country_level_poverty_data <- poverty %>% 
  filter(region != "Aggregates") %>% 
  filter(!is.na(SI.POV.DDAY)) 
```

Fix names to match Brauner list:

```{r}
country_level_poverty_data <- country_level_poverty_data %>% 
  mutate(
    country = case_when(
      country == "Slovak Republic" ~ "Slovakia",
      TRUE ~ country
    )
  )
```

What is the most recent year of data available by country?

```{r}
countries_by_year_of_most_recent_data <- country_level_poverty_data %>% 
  group_by(country) %>% 
  summarize(most_recent_year = max(year), .groups = "drop") %>% 
  arrange(desc(most_recent_year), country)
```


```{r}
countries_by_year_of_most_recent_data %>% 
  count(most_recent_year) %>% 
  arrange(desc(most_recent_year)) %>% 
  mutate(cumulative_n = cumsum(n))
```

Looking at the table above, if we included countries with data from 2016 or 
more recent, that would give us 101 countries.

```{r}
countries_with_poverty_data <- countries_by_year_of_most_recent_data %>% 
  filter(most_recent_year >= 2016) %>% 
  pull(country) %>% 
  sort()
```


## Age structure

```{r}
age %>% 
  filter(AgeGrpStart >= 65 & Time %in% 2010:2020) %>% 
  rename(country = Location, year = Time) %>% 
  group_by(country, year) %>% 
  summarize(complete = !any(is.na(PopTotal)), .groups = "drop") %>% 
  group_by(country) %>% 
  summarize(most_recent_year = max(year), .groups = "drop") %>% 
  arrange(desc(most_recent_year), country)
```



# Child mortality



# Map the Brauner countries


```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
```

```{r}
world %>% 
  mutate(is_a_brauner_country = name_long %in% brauner_countries) %>% 
  count(is_a_brauner_country)
```


```{r}
ggplot(data = world) +
  geom_sf()
```

