---
title: "Which countries have poverty data?"
author: "Derek Burk"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

# Packages

```{r}
library(readr)
library(stringr)
library(DT)
library(dplyr)
```

# File paths

```{r}
poverty_data_path <- "../downloaded_data/poverty-world-bank-2021-07-15.csv"
```

# Brauner et al. countries

```{r}
brauner_countries <- c(
  "Albania", "Andorra", "Austria", "Belgium", "Bosnia and Herzegovina", 
  "Bulgaria", "Croatia", "Czech Republic", "Denmark", "Estonia", "Finland", 
  "France", "Georgia", "Germany", "Greece", "Hungary", "Iceland", "Ireland", 
  "Israel", "Italy", "Latvia", "Lithuania", "Malaysia", "Malta", "Mexico", 
  "Morocco", "Netherlands", "New Zealand", "Norway", "Poland", "Portugal", 
  "Romania", "Serbia", "Singapore", "Slovakia", "Slovenia", "South Africa", 
  "Spain", "Sweden", "Switzerland", "United Kingdom"
)
```


# Load data

```{r}
poverty_data <- read_csv(
  poverty_data_path,
  col_types = cols(
    iso2c = col_character(),
    country = col_character(),
    SI.POV.DDAY = col_double(),
    year = col_double(),
    iso3c = col_character(),
    region = col_character(),
    capital = col_character(),
    longitude = col_double(),
    latitude = col_double(),
    income = col_character(),
    lending = col_character()
  )
)
```

```{r}
country_level_poverty_data <- poverty_data %>% 
  filter(str_detect(iso2c, "^[A-Z]{2}$")) %>% 
  filter(!is.na(SI.POV.DDAY)) 
```


# Reconcile country names

```{r}
setdiff(
  brauner_countries, 
  country_level_poverty_data$country
)
```
It looks like Slovakia is just called "Slovak Republic" in the World Bank
poverty dataset, but Andorra, New Zealand, and Singapore don't have any
non-missing data in the poverty dataset. I'll rename Slovakia in the
poverty data for the sake of comparison:

```{r}
country_level_poverty_data <- country_level_poverty_data %>% 
  mutate(
    country = case_when(
      country == "Slovak Republic" ~ "Slovakia",
      TRUE ~ country
    )
  )
```


# Which countries have poverty data?

```{r}
countries_by_year_of_most_recent_data <- country_level_poverty_data %>% 
  group_by(country) %>% 
  summarize(most_recent_year = max(year), .groups = "drop") %>% 
  arrange(desc(most_recent_year), country)

countries_by_year_of_most_recent_data %>% datatable()
```


```{r}
countries_by_year_of_most_recent_data %>% 
  count(most_recent_year) %>% 
  arrange(desc(most_recent_year)) %>% 
  mutate(cumulative_n = cumsum(n))
```

What about Brauner countries specifically?


```{r}
brauner_countries_by_year_of_most_recent_data <- country_level_poverty_data %>% 
  filter(country %in% brauner_countries) %>% 
  group_by(country) %>% 
  summarize(most_recent_year = max(year), .groups = "drop") %>% 
  arrange(desc(most_recent_year), country)

brauner_countries_by_year_of_most_recent_data %>% datatable()
```


```{r}
brauner_countries_by_year_of_most_recent_data %>% 
  count(most_recent_year) %>% 
  arrange(desc(most_recent_year)) %>% 
  mutate(cumulative_n = cumsum(n))
```