---
title: "Mortality data by country and frequency"
author: "Derek Burk"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

# Packages

```{r}
library(readr)
library(dplyr)
```

# File paths

```{r}
wmd_data_path <- "../downloaded_data/wmd-2021-07-15.csv"
```

# Brauner et al. countries

```{r}
brauner_countries <- c(
  "Albania", "Andorra", "Austria", "Belgium", "Bosnia and Herzegovina", 
  "Bulgaria", "Croatia", "Czech Republic", "Denmark", "Estonia", "Finland", 
  "France", "Georgia", "Germany", "Greece", "Hungary", "Iceland", "Ireland", 
  "Israel", "Italy", "Latvia", "Lithuania", "Malaysia", "Malta", "Mexico", 
  "Morocco", "Netherlands", "New Zealand", "Norway", "Poland", "Portugal", 
  "Romania", "Serbia", "Singapore", "Slovakia", "Slovenia", "South Africa", 
  "Spain", "Sweden", "Switzerland", "United Kingdom"
)
```


# Load data

```{r}
wmd <- read_csv(
  wmd_data_path,
  col_types = cols(
    country_name = col_character(),
    year = col_double(),
    time = col_double(),
    time_unit = col_character(),
    deaths = col_double()
  )
)
```

# Reconcile country names

```{r}
setdiff(brauner_countries, wmd$country_name)
```
It looks like Bosnia and Herzegovina and Czech Republic are in the WMD dataset 
under different names, but Morocco does not appear to be in the WMD dataset.
I'll rename the countries in the WMD data for the sake of comparison:

```{r}
wmd <- wmd %>% 
  mutate(
    country_name = case_when(
      country_name == "Bosnia" ~ "Bosnia and Herzegovina",
      country_name == "Czechia" ~ "Czech Republic",
      TRUE ~ country_name
    )
  )
```



# All time units

```{r}
unique(wmd$time_unit)
```


# Countries with weekly data 2015-2020

```{r}
countries_with_complete_weekly_data <- wmd %>% 
  filter(time_unit == "weekly" & year %in% 2015:2020) %>% 
  group_by(country_name, year) %>% 
  summarize(has_all_weeks = all(1:52 %in% time), .groups = "drop") %>% 
  group_by(country_name) %>% 
  summarize(for_all_years = all(has_all_weeks), .groups = "drop") %>% 
  filter(for_all_years) %>% 
  pull(country_name) %>% 
  sort()
countries_with_complete_weekly_data
```

# Countries with monthly data 2015-2020

```{r}
countries_with_complete_monthly_data <- wmd %>% 
  filter(time_unit == "monthly" & year %in% 2015:2020) %>% 
  group_by(country_name, year) %>% 
  summarize(has_all_months = all(1:12 %in% time), .groups = "drop") %>% 
  group_by(country_name) %>% 
  summarize(for_all_years = all(has_all_months), .groups = "drop") %>% 
  filter(for_all_years) %>% 
  pull(country_name) %>% 
  sort()
countries_with_complete_monthly_data
```

# Countries with quarterly data 2015-2020

```{r}
countries_with_complete_quarterly_data <- wmd %>% 
  filter(time_unit %in% c("quarterly", "quarterly (Solar Hijri)") & 
           year %in% 2015:2020) %>% 
  group_by(country_name, year) %>% 
  summarize(has_all_quarters = all(1:4 %in% time), .groups = "drop") %>% 
  group_by(country_name) %>% 
  summarize(for_all_years = all(has_all_quarters), .groups = "drop") %>% 
  filter(for_all_years) %>% 
  pull(country_name) %>% 
  sort()
countries_with_complete_quarterly_data
```

# Countries that don't have complete weekly, monthly, or quarterly data

```{r}
countries_with_incomplete_data <- setdiff(
  unique(wmd$country_name), 
  unique(
    c(
      countries_with_complete_weekly_data, 
      countries_with_complete_monthly_data,
      countries_with_complete_quarterly_data
    )
  )
)
countries_with_incomplete_data
```

# Countries with complete monthly or weekly data

```{r}
countries_with_complete_data <- 
  c(countries_with_complete_weekly_data, countries_with_complete_monthly_data) %>% 
  unique() %>% 
  sort()
countries_with_complete_data
```

Do all the Brauner et al. countries have complete monthly or weekly mortality 
data?

```{r}
setdiff(brauner_countries, countries_with_complete_data)
```

Yes, with the exception of Morocco, which isn't part of the WMD.
